---
title: 'Assignment #3'
output:
  html_document:
    df_print: paged
---
***
<center> 
### Embeding a Demand Function into an ARIMA Model:
### Dynamic Regression Laboratory.
#### (Total 40 pts.)
#### Due: Dec. 2 (before 9:00am)
</center>
***

The sales data is provided in CSV format in the file **"Peanut Butter Chicago.csv"**. As this is an individual skill-building assignment (as opposed to an open-ended team assignment), and I would like to achieve some degree of convergence in your answers, hence I have provided a common data preprocessing script below.  The data set corresponds to the total weekly sales of peanut butter for a supermarket chain, not for the individual stores. As you can observe from the file, the data corresponds to a combination of multiple brands as well as the supermarket private label (generic) in sizes ranging from 0.75 to 1.5 lbs.   

The data includes the following information for each individual stock keeping unit (SKU) as identified by its UPC code on each week in the data file:

* VEND	Number identifying the product vendor (48001 corresponds to Unilever).
* UPC	The product’s universal product code (bar code)
* UNITS	Sales volume
* DOLLARS	Dollar sales revenue
*	VOL_EQ	Weight in pounds of a units sold
*	PPU	Price per unit ($/lb)
*	F	Factor specifying advertising in the store weekly flyer:  
    + F = “A”	Large size ad.
    + F = “B”	Medium size ad.
    + F = “C”	Small size ad.
*	D	Factor specifying In-Store Display
    + D = 0	No In-Store Display
    + D = 1	Minor In-Store Display
    + D = 2	Major In-Store Display

To simplify the modeling process (and the assignment) in the preprocessing script  below I lumped all products into just three aggregate products (sub-categories): “SK” includes all Skippy brand products, “OB” includes all other branded products and “PL” includes all private label products. For each of the three aggregate products I obtained the total sales (volume) in pounds, the average sale prices ($/lb), and volume-weighted averages of the advertising and display variables (F and D).  Please take a few minutes to examine the pre-processing script below. 

 
Our goal is to embed a log-log demand model in an ARIMA model (Regression with ARIMA errors) that accounts for the auto-correlations in the sales data.  As a first attempt we would like to include a demand function of the following form:

$$y=e^{βx} p_S^α p_B^γ p_P^{γ_o}$$

Where the model variables and parameters are defined as follows:

* $y$ :	Demand (sales volume)
* $p_S$ :	Average price per pound of “Skippy” products
* $p_B$ :	Average price per pound of “Other Branded” products
* $p_P$ :	Average price per pound of “Private Label” products
* $x$ :	Vector of weighted averages of advertising and display variables for each product sub-category
* $β$ :	Vector of coefficients for advertising and display variables
* $α,γ,γ_o$:	Coefficients (elasticity and cross-elasticities) of prices

We have a total of 104 weeks of data.  In this assignment we will use weeks 1 through 94 as a training set and weeks 95 through 104 as a testing set.


```{r, message=FALSE, warning=FALSE}
library(fpp)
library(reshape)
library(dplyr)
library(glmnet)

# Data Pre-Processing 
#
PBS <- read.csv("PBS.csv") %>% 
  mutate( F_LSA=ifelse(F=="A",1,0),   # Large Size Ad Dummy
          F_MSA=ifelse(F=="B",1,0),   # Medium Size Ad Dummy
          F_SSA=ifelse(F=="C",1,0),   # Small Size Ad Dummy
          D_MIN=ifelse(D==1,1,0),     # Minor In-Store Display Dummy
          D_MAJ=ifelse(D==2,1,0)) %>% # Major In-Store Display Dummy
  # Promotional variables are weighted by sales volume (oz)
  mutate(S_LB = UNITS * VOL_EQ,
         WF_LSA = F_LSA * S_LB,     # Large Size Ad Weighted
         WF_MSA = F_MSA * S_LB,     # Medium Size Ad Weighted
         WF_SSA = F_SSA * S_LB,     # Small Size Ad Weighted
         WD_MIN = D_MIN * S_LB,     # Minor In-Store Display Weighted
         WD_MAJ = D_MAJ * S_LB) %>% # Major In-Store Display Weighted
  mutate(VEND =ifelse(VEND == 48001,"SK",ifelse( VEND == 99998,"PL","OB"))) %>%
  select(-F, -D)

# Create aggregate variables by product-week
x.pw <- PBS %>% group_by(WEEK, VEND) %>% 
  summarise(S.DOLLARS = sum(DOLLARS),      # Total $ Sales 
            S.S_LB    = sum(S_LB),         # Total L. Sales
            S.WF_LSA  = sum(WF_LSA),       # Total Weighted Large Ad
            S.WF_MSA  = sum(WF_MSA),       # Total Weighted Medium Ad
            S.WF_SSA  = sum(WF_SSA),       # Total Weighted Small Ad
            S.WD_MIN  = sum(WD_MIN),       # Total Weighted Minor Store Disp
            S.WD_MAJ  = sum(WD_MAJ)) %>%   # Total Weighted Major Store Disp
  # Calculate weigted averages of Advertising and Promotion variables
  mutate(A.PPU = log(S.DOLLARS / S.S_LB),  # Log of Avg. Price ($/pound)
         S.WF_LSA  = S.WF_LSA / S.S_LB,    # Avg. Weighted Large Ad
         S.WF_MSA  = S.WF_MSA / S.S_LB,    # Avg. Weighted Medium Ad
         S.WF_SSA  = S.WF_SSA / S.S_LB,    # Avg. Weighted Small Ad
         S.WD_MIN  = S.WD_MIN / S.S_LB,    # Avg. Weighted Minor Store Disp
         S.WD_MAJ  = S.WD_MAJ / S.S_LB)    # Avg. Weighted Major Store Disp

#
x.pw <- x.pw %>%
  mutate(LS  = log(S.S_LB)) %>% 
  select(-S.DOLLARS, -S.S_LB)
#
# Creeate separate dataframes for each brand group
x.SK <- x.pw %>% filter(VEND == "SK") %>% select(-VEND)
colnames(x.SK) <- c("WEEK","WF_LSA.SK","WF_MSA.SK","WF_SSA.SK","S.WD_MIN.SK","S.WD_MAJ.SK","PPU.SK","LS.SK" )
x.OB <- x.pw %>% filter(VEND == "OB") %>% select(-VEND,-LS)
colnames(x.OB) <- c("WEEK","WF_LSA.OB","WF_MSA.OB","WF_SSA.OB","S.WD_MIN.OB","S.WD_MAJ.OB","PPU.OB")
x.PL <- x.pw %>% filter(VEND == "PL") %>% select(-VEND,-LS)
colnames(x.PL) <- c("WEEK","WF_LSA.PL","WF_MSA.PL","WF_SSA.PL","S.WD_MIN.PL","S.WD_MAJ.PL","PPU.PL")

#Join the product-specific dataframes to create an expanded dataframe for SK using the 
# data from competing products as additional columns to be used as predicitve variables

xmat <- x.SK %>%
  left_join(x.OB,by="WEEK") %>%
  left_join(x.PL,by="WEEK")

# If your code executed correctly xmat should have 20 cols and 104 rows.
# 
xm <- model.matrix(LS.SK ~. - WEEK, data=xmat)[,-1]
y <- xmat$LS.SK

#Separation of Training and Testing sets
xm.tr <- xm[1:94,]
y.tr <-  y[1:94]
xm.te <- xm[95:104,]
y.te <-  y[95:104]
#
```


1. (5 pts) After pre-processing the data, notice that you have 18 predictive variables plus the week index and the sales vector.  Notice that the pre-processing step already computes the log of the average prices and sales volumes. Now use The Lasso on the training set to obtain (a) a regularized model and (b) the reduced set of predictive variables that minimize the cross-validated MSE over the training set (i.e., the set of variables included in the Lasso-regularized model). (Use set.seed(1) before 10-fold cross-validation).  Report the coefficients of the regularized model.
```{r}
set.seed(1)
CV.L = cv.glmnet(xm.tr, y.tr,alpha=1, nfolds=10)
LamL = CV.L$lambda.1se
coef.L = predict(CV.L,type="coefficients",s=LamL)
coef.L
```


2. (5 pts) Use the training set to fit an unrestricted regression model (i.e., **lm(…)** ) on the reduced set of explanatory variables identified by The Lasso.  Report the coefficients of the full model and comment on the fit of the model and examine the auto-correlations of the residuals of this model. 
```{r}
df2 <- cbind(y.tr, xm.tr[,'S.WD_MIN.SK'], xm.tr[,'PPU.SK'])
colnames(df2) <- c('LS.SK', 'S.WD_MIN.SK', 'PPU.SK')
df2 <- as.data.frame(df2)
lr <- lm(LS.SK~S.WD_MIN.SK+PPU.SK, data=df2)
summary(lr)
```


```{r}
checkresiduals(lr)
```


<span style="color:blue">
From the graph, there seems to be some autocorrelation in the residuals. Also H0 is rejected in the test for autocorrelation, so there is some serial correlation in the residuals. 
</span>


3. (5 pts) Fit a simple ARIMA model (not a dynamic regression model) to explain the training set log-of-sales-volume data. Report the diagnostic of your model’s residuals and comment on the model’s validity.  
```{r}
y.ts <- as.ts(y.tr)
L <- BoxCox.lambda(y.ts)
z <- y.ts %>% BoxCox(L)
fit.G <- z %>% auto.arima(stepwise=FALSE, approximation=FALSE)
checkresiduals(fit.G)
```

```{r}
summary(fit.G)
```


<span style="color:blue">
While the ARIMA model seems valid because errors are white noise and there does not seem to be autocorrelation, we need to check the performance on test data to see if it is actually usefull for forecasting.
</span>


4. (5 pts) Use the model in Question 3 to prepare a 10 period ahead forecast and compare it (overly it) with the testing set log-of-sales data.  Comment on the usefulness of this model in terms of precision and confidence interval. 

```{r}
autoplot(forecast(fit.G, h=10)) +
  autolayer(y.te %>% as.ts() %>% BoxCox(L) %>% stats::lag(-95), series = 'Test Data')
```


<span style="color:blue">
The model is not useful for forecasting because the confidence intervals are too large. The model also seems to be biased downwards.
</span>


5. (5 pts) Use the **auto.arima(…)** function to fit a dynamic regression (i.e., regression with ARIMA errors) model to explain sales data (log) using only the predictive variables identified by The Lasso in Question 1.  Examine the model’s residuals and comment on its validity. 


```{r}
xreg <- cbind(S.WD_MIN.SK = df2$S.WD_MIN.SK,
              PPU.SK = df2$PPU.SK)
fit.G2 <- auto.arima(z, xreg=xreg)
summary(fit.G2)
```

```{r}
checkresiduals(fit.G2)
```



<span style="color:blue">
The dynamic regression model seems valid because errors are white noise and there does not seem to be autocorrelation.
</span>


6. (5 pts) Obtain a dynamic regression model that improves on the auto-arima model in Question 5 in terms of its information coefficients and residual diagnostics. Compare the coefficients of the explanatory variables in (a) The Lasso model, (b) The unrestricted model obtained in Question 2, and (c) The ones obtained in this question.  Then use the B notation (polynomial) to describe the model you obtained. 

```{r}
xreg2 <- cbind(S.WD_MIN.SK = df2$S.WD_MIN.SK,
               PPU.SK = df2$PPU.SK)

fit.G3 <- auto.arima(z, xreg=xreg2, stepwise=FALSE, approximation=FALSE)
summary(fit.G3)
```

```{r}
checkresiduals(fit.G3)
```

<span style="color:blue">
Lasso: Coefficients are closer to 0 when compared to the unrestricted model because of regulation <br/>
Unrestricted: Coefficients are largest of the three <br/>
Dynamic regression: Coefficients are lower because there is an additional ARIMA model explaining part of the variation <br/>
</span>
B Notaion: <br/>

$y'_t = \beta_1 x'_{1,t} +  \beta_2 x'_{2,t} + n'_t$ <br/>
where <br/>
$(1-\phi_1B)n'_t = (1+\theta_1B+\theta_2B)e_t$ <br/>


7. (5 pts) Use the model in Question 5 to prepare a 10 period ahead forecast and compare it (overlay it) with the testing set log-of-sales data. You can also obtain the values of the regressors used in the forecasting model from the testing data set **xm.te**.  Comment on the usefulness of this model in terms of precision and confidence interval relative to the model without explanatory variables in Question 3.  
```{r}
df7 <- cbind(xm.te[,'S.WD_MIN.SK'], xm.te[,'PPU.SK'])
colnames(df7) <- c('S.WD_MIN.SK', 'PPU.SK')
autoplot(forecast(fit.G2, xreg=df7, h=10)) +
  autolayer(y.te %>% as.ts() %>% BoxCox(L) %>% stats::lag(-95), series = 'Test Data')
```


<span style="color:blue">
The model is more useful for forecasting now compared to before, because the confidence interval is narrower now. The model still seems to be biased downward though.
</span>



8. (5 pts) After you complete a project, it is often useful to reflect on what would you do different if you were to perform this project again.  This is no exception.  Comment on the training and testing fit statistics and discuss how do you think you could improve on the performance of the model in terms of (a) additional data, (b) different pre-processing of the existing data, and (c) different modeling choices.  Discuss your assessment of the potential for improvement (ex-ante priorities) for the different improvement options you suggest.
<br/>

<span style="color:blue">
a.) I would consider adding total sales in competing retail chains, if it is available, to model diminishing demands in case of increased competition. Also add variables scraped from weather reports, becuase customers tend to stock up on long-lasting food such as peanut butter when they anticipate hurricanes or heavy snow. 
<br/>
b.) Based on coefficients from the Lasso regression, the effect of demand changes in the Skippy brand seems to dominate all the other brands. So we should probably model the demand for each type of peanut butters individually.
<br/>
c.) Due to the large amount of features, I may try tree-based ensamble regression models with dummy variables for seasonality and special promotions.
</span>
